---
title: "RollingSales"
author: "Andrew Abbott, Ken Avery, Paola Leon, Earl Shaw"
date: "May 30, 2016"
output:
        html_document: 
          keep_md: true
---

```{r}
setwd("/Users/Abbott/test-repo/RollingSales")
```

# Save the file as a csv and use read.csv
```{r}
queens <- read.csv("rollingsales_queens.csv",skip=4,header=TRUE)
```

## Check the data
```{r}
head(queens)
summary(queens)
str(queens)
```

# We create a new variable that is a "clean" version of sale.price.
# sale.price.n is numeric, not a factor.
```{r}
library(plyr)
queens$SALE.PRICE.N <- as.numeric(gsub("[^[:digit:]]","", queens$SALE.PRICE))
count(is.na(queens$SALE.PRICE.N))
```
There are 7,488 without values.

# Make all variable names lower case
```{r}
names(queens) <- tolower(names(queens)) 
names(queens)
```

## Get rid of leading digits and make sure numeric

```{r}
queens$gross.sqft <- as.numeric(gsub("[^[:digit:]]","", queens$gross.square.feet))
queens$land.sqft <- as.numeric(gsub("[^[:digit:]]","", queens$land.square.feet))
class(queens$gross.sqft)
class(queens$land.sqft)
```

## Change sale.date to Date format in R, format must be specified.
```{r}
queens$sale.date <- as.Date(queens$sale.date, "%m/%d/%y")
queens$year.built <- as.numeric(as.character(queens$year.built))
class(queens$sale.date)
```


## Do a bit of exploration to make sure there's not anything
## weird going on with sale prices
```{r}
attach(queens)
hist(sale.price.n) 
```

# Everything looks to be 0
```{r}
hist(sale.price.n[sale.price.n>0])
```

#That didn't help
```{r}
hist(sale.price.n[sale.price.n<1000000])
```

#that looks reasonable

```{r}
detach(queens)
```

##Now that the data is in correct formats and types we will write it to a new .csv file.
```{r}
write.csv(queens, file = "rollingsales_queens2.csv", )
```



## Lets keep only the actual sales and create a new data frame to hold them.
##First lets load the clean data.

```{r}
queens2 <- read.csv("rollingsales_queens2.csv",header=TRUE)
```

##Next lets filter out sales prices of $0.00

```{r}
queens.sale <-queens2[queens2$sale.price.n!=0,]
```

## We include only sales with prices greater than $0.

```{r}
plot(queens.sale$gross.sqft,queens.sale$sale.price.n)
```
```{r}
plot(log(queens.sale$gross.sqft),log(queens.sale$sale.price.n))
```

##Save this data step in a csv file.

```{r}
write.csv(queens.sale, file = "rollingsales_queenssale.csv")
```

## There are still some really large sales prices muddling the data.
## for now, let's look at 1-, 2-, and 3-family homes only
## and create a new data frame to hold them.

```{r}
queens3 <- read.csv("rollingsales_queenssale.csv",header=TRUE)
```


```{r}
queens.homes <- queens3[which(grepl("FAMILY",queens3$building.class.category)),]
dim(queens.homes)
```

```{r}
hist(queens.homes$sale.price.n)
```
```{r}
plot(log(queens.homes$gross.sqft),log(queens.homes$sale.price.n))
summary(queens.homes[which(queens.homes$sale.price.n<100000),])
```

## Now to remove outliers that seem like they weren't actual sales
```{r}
queens.homes$outliers <- (log(queens.homes$sale.price.n) <=10) + 0
queens.homes <- queens.homes[which(queens.homes$outliers==0),]
plot(log(queens.homes$gross.sqft),log(queens.homes$sale.price.n))
```

## Finally save the dataframe as new csv files:

```{r}
write.csv(queens.homes, file = "rollingsales_queenshomes.csv")
```

